Measuring contrast processing in the visual system using SSVEP

### 2024-05-13 Rversion/reviewR.qmd
Added R version in subfolder

```{python setup}

# all package loading and subfunctions defined here
import numpy as np
import matplotlib.pyplot as plt
from matplotlib import colormaps
import scipy as scp
from sklearn.linear_model import LinearRegression

def simpleTransducerFunction(inputContrast, c50, RMax):
 # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
 # RMax defines the reponse gain of the transducer (how big the biggest output is). 
 # c50 defines the conrast sensitivity. Smaller values are more sensitive.
 # The exponent can also vary - here we fix it to 2
  expnt=2

  # Contrast cannot be negative. To model the responses of a set of on and off 
  # contrast detectors we perform full-wave rectification of the signal
  inputContrast=np.abs(inputContrast)
  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + c50**expnt)
  return output

def doubleTransducerFunction(inputContrast, maskContrast, c50, RMax):
 # This version has two inputs, but the second input is purely suppressive
  expnt=2
  inputContrast=np.abs(inputContrast)
  maskContrast=np.abs(maskContrast)

  output=RMax*(inputContrast**expnt)/(inputContrast**expnt + maskContrast**expnt + c50**expnt)
  return output

def create_grating(size, frequency, phase, contrast):
    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
    grating = np.sin(2 * np.pi * frequency * np.sqrt(x**2 + y**2) + phase)
    grating = (grating + 1) / 2  # Normalize to [0, 1]
    grating = grating * contrast + (1 - contrast) / 2  # Adjust contrast
    return grating

def create_circular_mask(size):
    x, y = np.meshgrid(np.linspace(-1, 1, size), np.linspace(-1, 1, size))
    mask = np.sqrt(x**2 + y**2) <= 0.3
    return mask

@@

```{python createcentresurround}

# Set parameters
size = 256
frequency = 10
phase = 0
contrast_center = 0.5
contrast_surround_low = 0.1
contrast_surround_high = 0.9

# Create gratings and circular masks
grating_center = create_grating(size, frequency, phase, contrast_center)
grating_surround_low = create_grating(size, frequency, phase, contrast_surround_low)
grating_surround_high = create_grating(size, frequency, phase, contrast_surround_high)
mask = create_circular_mask(size)

# Combine gratings and masks
image_low_contrast = grating_surround_low.copy()
image_low_contrast[mask] = grating_center[mask]
image_high_contrast = grating_surround_high.copy()
image_high_contrast[mask] = grating_center[mask]

# Create subplots
fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(7, 3))

# Plot low-contrast background
ax1.imshow(image_low_contrast, cmap='gray', vmin=0, vmax=1)
ax1.axis('off')
ax1.set_title('Low-Contrast Background', )

# Plot high-contrast background
ax2.imshow(image_high_contrast, cmap='gray', vmin=0, vmax=1)
ax2.axis('off')
ax2.set_title('High-Contrast Background')

plt.tight_layout()
plt.savefig('Figures/centresurround.pdf')

@@

```{python linearsystem}

inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30
c50=.10
RMax=10

support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)

initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal=[]

initialSignal = np.zeros(len(contrastRange)) 
transducedSignal = np.zeros(len(contrastRange))
for contrastIndex,thisContrast in enumerate(contrastRange):
  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast

  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)

  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
  initialSignal[contrastIndex]=inputAmplitude[0]

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots
fig, axs = plt.subplots(1,3, figsize=(8, 3))

axs[0].plot(support_S, inputModulation)
axs[0].set_title('Input sine wave', fontsize=10)
axs[0].set_xlabel('Time (s)')

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_Signal[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange*100,initialSignal)
axs[2].set_xlabel('Contrast (%)')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Linear computation stage', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plt.savefig('Figures/linearsystem.pdf')

@@

```{python transducerout}

# Define some constants
inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30
c50=.10
RMax=10

support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)

initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal=[]

initialSignal = np.zeros(len(contrastRange)) 
transducedSignal = np.zeros(len(contrastRange))
for contrastIndex,thisContrast in enumerate(contrastRange):
  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast

  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)

  inputAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
  initialSignal[contrastIndex]=inputAmplitude[0]

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots
fig, axs = plt.subplots(1,3, figsize=(8, 3))

axs[0].plot(support_S, transducedModulation)
axs[0].set_title('Transduced signal', fontsize=10)
axs[0].set_xlabel('Time (s)')

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange*100,transducedSignal)
axs[2].set_xlabel('Contrast (%)')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Transducer output', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plt.savefig('Figures/transducer.pdf')

@@

```{python twoinputsout}

# Define some constants
inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
maskFrequency_Hz=np.array([7]) # We simulate a system with on input at many different contrasts
maskContrast=np.max(contrastRange) # contrast of mask signal

duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30
c50=.10
RMax=10

support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)

initialSignal=[] # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal=[]

initialSignal = np.zeros(len(contrastRange)) 
transducedSignal = np.zeros(len(contrastRange))
maskedSignal = np.zeros(len(contrastRange))
for contrastIndex,thisContrast in enumerate(contrastRange):
  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
  maskModulation = np.sin(2 * np.pi * maskFrequency_Hz * support_S) * maskContrast

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

  transducedModulation=doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  maskedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots
fig, axs = plt.subplots(1,3, figsize=(8, 3))

axs[0].plot(support_S, transducedModulation)
axs[0].set_title('Transduced signal', fontsize=10)
axs[0].set_xlabel('Time (s)')

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange*100,transducedSignal)
axs[2].plot(contrastRange*100,maskedSignal)
axs[2].set_xlabel('Contrast (%)')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Transducer output', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plt.savefig('Figures/twoinputs.pdf')

@@

```{python makesweep}

# Define some constants
nLevels=8
internalNoise=1
inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(1,1.9,nLevels)/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30

support_S=np.linspace(0,duration_S,duration_S*digitisationRate_Hz)

fig, axs = plt.subplots(1,1, figsize=(4, 3))

initialSignal = np.zeros(len(contrastRange))
transducedSignal = np.zeros(len(contrastRange))
for contrastIndex,thisContrast in enumerate(contrastRange):
  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast + np.random.rand(len(support_S))*internalNoise

  FT_Signal = np.fft.fft(inputModulation) / len(inputModulation)

  responseAmplitude = np.abs(FT_Signal[inputFrequency_Hz])
  initialSignal[contrastIndex]=responseAmplitude[0]
  transducedSignal[contrastIndex]=simpleTransducerFunction(responseAmplitude[0], 0.05, 10)

# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
X = np.log10(contrastRange[:4]).reshape(-1, 1)
X2 = np.log10(np.logspace(0.1,1.1,nLevels)/100).reshape(-1, 1)
y = transducedSignal[:4]
reg = LinearRegression().fit(X, y)
x_intercept = 10 ** (-reg.intercept_ / reg.coef_[0])

axs.set_title(f'Simulated sweep VEP at {nLevels} different contrasts', fontsize=10)
axs.semilogx(contrastRange, transducedSignal, 'o')
axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')
axs.semilogx([x_intercept], [0], 'ro', label='Threshold')
axs.set_xlabel('Contrast')
axs.set_ylabel('Amplitude (a.u.)')
axs.set_xlim(0.01, 1)
axs.set_ylim(0, 10)
axs.legend()
axs.grid()
plt.tight_layout()
plt.savefig('Figures/sweepplot.pdf')

@@

```{python makeclinicalplot}

# Parameters
finecontrastsdB = np.arange(-18, 37, 1)
finecontrasts = 10**(finecontrastsdB / 20)
ticklocsx = np.arange(-18, 37, 6)
ticklabelsx = ['1/8', '1/4', '1/2', 1, 2, 4, 8, 16, 32, 64]

epilepsyparams = np.array([
    [3.4, 0.15, 1.4, 1],  # controls
    [5.3, 0.23, 1.4, 1]   # patients
])

# Calculate responses
respEPIcontrols = epilepsyparams[0,0] * ((finecontrasts/100)**epilepsyparams[0,2] /
                (epilepsyparams[0,1]**epilepsyparams[0,2] + (finecontrasts/100)**epilepsyparams[0,2])) + epilepsyparams[0,3]

respEPIpatients = epilepsyparams[1,0] * ((finecontrasts/100)**epilepsyparams[1,2] /
                (epilepsyparams[1,1]**epilepsyparams[1,2] + (finecontrasts/100)**epilepsyparams[1,2])) + epilepsyparams[1,3]

>>>>

```{r setup}

# set up helper functions

simpleTransducerFunction <- function(inputContrast, c50, RMax){
 # Here we model a simple hyperbolic ratio function. Also called a "Naka-Rushton" function.
 # RMax defines the reponse gain of the transducer (how big the biggest output is). 
 # c50 defines the conrast sensitivity. Smaller values are more sensitive.
 # The exponent can also vary - here we fix it to 2
  expnt <- 2

  # Contrast cannot be negative. To model the responses of a set of on and off 
  # contrast detectors we perform full-wave rectification of the signal
  inputContrast <- abs(inputContrast)
  output <- RMax*(inputContrast^expnt)/(inputContrast^expnt + c50^expnt)
  return(output)
}

doubleTransducerFunction <- function(inputContrast, maskContrast, c50, RMax){
 # This version has two inputs, but the second input is purely suppressive
  expnt <- 2
  inputContrast <- abs(inputContrast)
  maskContrast <- abs(maskContrast)

  output <- RMax*(inputContrast^expnt)/(inputContrast^expnt + maskContrast^expnt + c50^expnt)
  return(output)  
}

create_grating <- function(size, frequency, phase, contrast){
    x <- matrix(rep(seq(-1,1,length=size),size),nrow=size,ncol=size)
    y <- t(x)
    grating <- sin(2 * pi * frequency * sqrt(x^2 + y^2) + phase)
    grating <- (grating + 1) / 2  # Normalize to [0, 1]
    grating <- grating * contrast + (1 - contrast) / 2  # Adjust contrast 
  return(grating)
}

create_circular_mask <- function(size){
    x <- matrix(rep(seq(-1,1,length=size),size),nrow=size,ncol=size)
    y <- t(x)
    radius <- sqrt(x^2 + y^2)
    mask <- sqrt(x**2 + y**2) <= 0.3
  return(mask)
}

if (!dir.exists('/Figures')){dir.create('/Figures')}

@@

```{r createcentresurround}

# Set parameters
size <- 256
frequency <- 10
phase <- 0
contrast_center <- 0.5
contrast_surround_low <- 0.1
contrast_surround_high <- 0.9

# Create gratings and circular masks
grating_center <- create_grating(size, frequency, phase, contrast_center)
grating_surround_low <- create_grating(size, frequency, phase, contrast_surround_low)
grating_surround_high <- create_grating(size, frequency, phase, contrast_surround_high)
mask <- create_circular_mask(size)

# Combine gratings and masks
image_low_contrast <- grating_surround_low
image_low_contrast[mask] <- grating_center[mask]
image_high_contrast <- grating_surround_high
image_high_contrast[mask] <- grating_center[mask]

pdf('Figures/centresurround.pdf',width=12,height=6.5)

par(mfcol=c(1,2))

plot(x=NULL,y=NULL,xlim=c(0,1),ylim=c(0,1),axes=FALSE, ann=FALSE)
image(image_low_contrast,col=gray.colors(256),add=TRUE,useRaster=TRUE)
title(main='Low-Contrast Background')

plot(x=NULL,y=NULL,xlim=c(0,1),ylim=c(0,1),axes=FALSE, ann=FALSE)
image(image_high_contrast,col=gray.colors(256),add=TRUE,useRaster=TRUE)
title(main='High-Contrast Background')

dev.off()

@@

```{r linearsystem}

inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S <- 1 # All variables in this code have an _[units] where possible
digitisationRate_Hz <- 1000
freqCutoff_Hz <- 30
c50 <- 0.10
RMax <- 10

support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)

initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal <- NULL

for (contrastIndex in 1:length(contrastRange)){

  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]

  FT_Signal <- fft(inputModulation) / length(inputModulation)

  inputAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
  initialSignal[contrastIndex] <- inputAmplitude

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation <- simpleTransducerFunction(inputModulation, c50, RMax)
  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
  transducedAmplitude <- abs(FT_transducedModulation[1 + inputFrequency_Hz*2])
  transducedSignal[contrastIndex] <- transducedAmplitude
}

frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots

pdf('Figures/linearsystem.pdf',width=12,height=4)
par(mfrow=c(1,3))

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(-0.3,0.3))   
  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(-3,3,1)/10, tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
  mtext(text = seq(-3,3,1)/10, side = 2, at=seq(-3,3,1)/10, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(main="Input sine wave", cex.main=1.5)

lines(support_S, inputModulation, lwd=3)

# Plot the absolute values of the FFT in the second axis as a bar plot

  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_Signal[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,0.16))
  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,0.16,0.04), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
  mtext(text = seq(0,0.16,0.04), side = 2, at=seq(0,0.16,0.04), line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Fourier spectrum", cex.main=1.5)

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,0.16))   
  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,0.16,0.04), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
  mtext(text = seq(0,0.16,0.04), side = 2, at=seq(0,0.16,0.04), line=0.2, las=1)  
title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Linear computation stage", cex.main=1.5)

lines(contrastRange*100,initialSignal,lwd=3)

dev.off()

@@

```{r transducerout}

inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S <- 1 # All variables in this code have an _[units] where possible
digitisationRate_Hz <- 1000
freqCutoff_Hz <- 30
c50 <- 0.10
RMax <- 10

support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)

initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal <- NULL

for (contrastIndex in 1:length(contrastRange)){

  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]

  FT_Signal <- fft(inputModulation) / length(inputModulation)

  inputAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
  initialSignal[contrastIndex] <- inputAmplitude

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation <- simpleTransducerFunction(inputModulation, c50, RMax)
  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
  transducedAmplitude <- abs(FT_transducedModulation[1 + inputFrequency_Hz*2])
  transducedSignal[contrastIndex] <- transducedAmplitude
}

frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots

pdf('Figures/transducer.pdf',width=12,height=4)
par(mfrow=c(1,3))

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,10))   
  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(main="Input sine wave", cex.main=1.5)

lines(support_S, transducedModulation, lwd=3)

# Plot the absolute values of the FFT in the second axis as a bar plot

  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_transducedModulation[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,2))
  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Fourier spectrum", cex.main=1.5)

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,2))   
  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Linear computation stage", cex.main=1.5)

lines(contrastRange*100,transducedSignal,lwd=3)

dev.off()

@@

```{r twoinputsout}

inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
contrastRange <- (10^seq(0,1.5,length=20))/100 # Equally spaced log contrast between 1 and 10^1.5
maskFrequency_Hz <- 7
maskContrast <- max(contrastRange)

duration_S <- 1 # All variables in this code have an _[units] where possible
digitisationRate_Hz <- 1000
freqCutoff_Hz <- 30
c50 <- 0.10
RMax <- 10

support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)

initialSignal <- NULL # Create lists to hold the initial contrast representation and the result of passign it through a simple Naka-Rushton function
transducedSignal <- NULL
maskedSignal <- NULL

for (contrastIndex in 1:length(contrastRange)){

  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex]
  maskModulation <- sin(2 * pi * maskFrequency_Hz * support_S) * maskContrast

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation <- doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
  transducedAmplitude <- abs(FT_transducedModulation[1+inputFrequency_Hz*2])
  transducedSignal[contrastIndex] <- transducedAmplitude

  transducedModulation <- doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
  FT_transducedModulation <- fft(transducedModulation)/ length(transducedModulation)
  transducedAmplitude <- abs(FT_transducedModulation[1+inputFrequency_Hz*2])
  maskedSignal[contrastIndex] <- transducedAmplitude
}

frequencySupport <- seq(0,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

#Make a figure with three subplots

pdf('Figures/twoinputs.pdf',width=12,height=4)
par(mfrow=c(1,3))

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,10))   
  axis(1, at=seq(0,1,0.5), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,1,0.5), side = 1, at=seq(0,1,0.5), line=0.2)     # add the tick labels
  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(main="Input sine wave", cex.main=1.5)

lines(support_S, transducedModulation, lwd=3)

# Plot the absolute values of the FFT in the second axis as a bar plot

  plot(frequencySupport[2:length(frequencySupport)]-0.5,abs(FT_transducedModulation[2:length(frequencySupport)]),type='s',lwd=3,axes=FALSE, ann=FALSE, xlim=c(0,30), ylim=c(0,2))
  axis(1, at=seq(0,30,10), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,30,10), side = 1, at=seq(0,30,10), line=0.2)     # add the tick labels
  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Magnitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Fourier spectrum", cex.main=1.5)

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,32), ylim=c(0,2))   
  axis(1, at=seq(0,32,4), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,2,0.5), tck=0.01, lab=F, lwd=2)
  mtext(text = seq(0,32,4), side = 1, at=seq(0,32,4), line=0.2)     # add the tick labels
  mtext(text = seq(0,2,0.5), side = 2, at=seq(0,2,0.5), line=0.2, las=1)  
title(xlab="Contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=2.6, cex.lab=1.5)
title(main="Linear computation stage", cex.main=1.5)

lines(contrastRange*100,transducedSignal,lwd=3)
lines(contrastRange*100,maskedSignal,col='orange',lwd=3)

dev.off()

@@

```{r makesweep}

# Define some constants
nLevels <- 8
internalNoise <- 1
inputFrequency_Hz <- 5 # We simulate a system with on input at many different contrasts
contrastRange <- (10^seq(1,1.9,length=nLevels))/100  # Equally spaced log contrast between 1 and 10^1.9
duration_S <- 1 # All variables in this code have an _[units] where possible
digitisationRate_Hz <- 1000
freqCutoff_Hz <- 30

support_S <- seq(0,duration_S,length=duration_S*digitisationRate_Hz)

initialSignal <- NULL
transducedSignal <- NULL
for (contrastIndex in 1:length(contrastRange)){
  inputModulation <- sin(2 * pi * inputFrequency_Hz * support_S) * contrastRange[contrastIndex] + runif(length(support_S))*internalNoise

  FT_Signal <- fft(inputModulation) / length(inputModulation)

  responseAmplitude <- abs(FT_Signal[inputFrequency_Hz+1])
  initialSignal[contrastIndex] <- responseAmplitude
  transducedSignal[contrastIndex] <- simpleTransducerFunction(responseAmplitude, 0.05, 10)
}
# Now fit a linear regression to the first 4 contrast levels (before it saturates) and find out where this crosses the Y-axis
X = log10(contrastRange[1:4])
X2 = log10((10^seq(0.1,1.1,length=nLevels))/100)
y <- transducedSignal[1:4]
reg <- lm(y ~ X)
x_intercept <- 10 ^ (-reg$coefficients[1] / reg$coefficients[2])
Yfit <- reg$coefficients[2]*X + reg$coefficients[1]
Yextrap <- reg$coefficients[2]*X2 + reg$coefficients[1]

pdf('Figures/sweepplot.pdf',width=6,height=6)

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-2,0), ylim=c(0,10))   
  axis(1, at=seq(-2,0,1), tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=seq(0,10,2), tck=0.01, lab=F, lwd=2)
  mtext(text = c(0.01,0.1,1), side = 1, at=seq(-2,0,1), line=0.2)     # add the tick labels
  mtext(text = seq(0,10,2), side = 2, at=seq(0,10,2), line=0.2, las=1)  
title(xlab="Contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(ylab="Amplitude (a.u.)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    
title(main=paste("Simulated sweep VEP at",nLevels,"different contrasts"), cex.main=1.5)

points(log10(contrastRange),transducedSignal,pch=19,col='cornflowerblue',cex=2)
points(log10(x_intercept),0,pch=19,col='red',cex=2)

lines(X,Yfit,lwd=3,col='red')
lines(X2,Yextrap,lwd=3,lty=2,col='red')

legend(-2,10,c('Linear fit','Threshold'),lty=c(1,NA),pch=c(NA,19),lwd=3,col='red',box.lwd=2,cex=1.5)

dev.off()

# axs.semilogx(contrastRange[:4], reg.predict(X), 'r-', label='Linear Fit')
# axs.semilogx(np.logspace(0.1,1.1,nLevels)/100, reg.predict(X2), 'r--')

@@

```{r makeclinicalplot}

# Parameters
finecontrastsdB <- seq(-18,36,1)
finecontrasts <- 10^(finecontrastsdB / 20)
ticklocsx <- seq(-18,36,6)
ticklabelsx <- c('1/8', '1/4', '1/2', 1, 2, 4, 8, 16, 32, 64)

pdf('Figures/clinicalplot.pdf',width=12,height=12)
par(mfrow=c(2,2))

epilepsyparams <- matrix(0,nrow=2,ncol=4) 
epilepsyparams[1,] <- c(3.4, 0.15, 1.4, 1)  # controls
epilepsyparams[2,] <- c(5.3, 0.23, 1.4, 1)  # patients

# Calculate responses
respEPIcontrols <- epilepsyparams[1,1] * ((finecontrasts/100)^epilepsyparams[1,3] /
                (epilepsyparams[1,2]^epilepsyparams[1,3] + (finecontrasts/100)^epilepsyparams[1,3])) + epilepsyparams[1,4]

respEPIpatients <- epilepsyparams[2,1] * ((finecontrasts/100)^epilepsyparams[2,3] /
                (epilepsyparams[2,2]^epilepsyparams[2,3] + (finecontrasts/100)^epilepsyparams[2,3])) + epilepsyparams[2,4]
